---
title: "CKME 136 Capstone"
author: "TODD BETHELL"
date: "July 1, 2019"
output:
  word_document: default
---
install.packages("caret")
install.packages("PerformanceAnalytics")
install.packages("lsr")
install.packages("tidyverse")
install.packages("dplyr")
install.packages("class")
install.packages("gmodels")
install.packages("corrplot")
install.packages("RCurl")
install.packages("GGally")
install.packages("FSelector")
install.packages("FSelectorRcpp")
library(GGally)
library(knitr)
library(RCurl)
library(caret)
library(PerformanceAnalytics)
library(lsr)
library(tidyverse)
library(dplyr)
library(class)
library(gmodels)
library(corrplot)
library(GGally)
library(FSelectorRcpp)
---
```{r results='hide', message=FALSE, warning=FALSE}
library(knitr)
library(RCurl)
library(caret)
library(PerformanceAnalytics)
library(lsr)
library(tidyverse)
library(dplyr)
library(plyr)
library(class)
library(gmodels)
library(corrplot)
library(GGally)
library(FSelectorRcpp)
```


```{r}
FullSetB4<-read.csv(file="https://raw.githubusercontent.com/ToddB11/TB136capstone/master/ZAlizadeh_dataset.csv",header=T,sep=",")
```

# Review Data Set
```{r}
head(FullSetB4)
tail(FullSetB4)
str(FullSetB4)
```

# Fixing 'Age' column header (not displaying correctly):
```{r}
names(FullSetB4)[1] <- "Age"
```
# Remove 'Length' attribute - since not used in final study; and captured via BMI (and 'Obesity' uses 'BMI'):
```{r}
FullSetB4 <- FullSetB4[-3]
dim(FullSetB4)
```

#create an ID field for possible merging later:
```{r}
# FullB4wID <- FullSetB4

# FullB4wID$ID <- seq(dim(FullB4wID)[1])

# FullB4wID <- FullB4wID[,c(55, 1:54)]
# FullB4wID

```
# CHECKING FOR MISSING DATA & ERRORS:

# Check number of observations and whether there is any missing data:
```{r}
nrow(FullSetB4)
sum(is.na(FullSetB4))
```

# Double checking whether all observations are complete (no missing data):  No incomplete cases found, but I am commenting this out for the html output - since the output is very long.
```{r}
#FullSetB4[complete.cases(FullSetB4),]
```

# Checking for errors, starting with 'Sex' attribute (column 3):
```{r}
SexF <- sum(FullSetB4$Sex == 'Fmale')
SexM <- sum(FullSetB4$Sex == 'Male')
sum(SexF + SexM)
```

# Checking for errors in columns 5:9, 19, 24, 31:34 (must be '0' or '1')
```{r}
#library(tidyverse)
#library(dbplyr)

Bins0 <- FullSetB4%>%
  gather(x, value, 5:9,19,24,31:34)%>%
  tally(value == 0)
Bins1 <- FullSetB4%>%
  gather(x, value, 5:9,19,24,31:34)%>%
  tally(value == 1)
 (Bins0 + Bins1)/11
```
# Checking for errors in columns 10:16, 20:23, 25, 27:28, 30, 35:36  (must be 'Y' or 'N')
```{r}

BinsY <- FullSetB4%>%
  gather(x, value, 10:16,20:23,25,27:28,30,35:36)%>%
  tally(value == 'Y')
BinsN <- FullSetB4%>%
  gather(x, value, 10:16,20:23,25,27:28,30,35:36)%>%
  tally(value == 'N')
 (BinsY + BinsN)/17
```
# Checking for errors in the 'BBB'(Bundle Branch Block) attribute (column 37):
```{r}
nBBB <- sum(FullSetB4$BBB == 'N')
LBBB <- sum(FullSetB4$BBB == 'LBBB')
RBBB <- sum(FullSetB4$BBB == 'RBBB')
sum(nBBB+LBBB+RBBB)
```
# Checking for errors in the 'VHD'(Valvular Heart Disease) attribute (column 54):
```{r}
Vmild <- sum(FullSetB4$VHD == 'mild')
VMod <- sum(FullSetB4$VHD == 'Moderate')
VN <- sum(FullSetB4$VHD == 'N')
VSev <- sum(FullSetB4$VHD == 'Severe')
sum(Vmild+VMod+VN+VSev)
```

# Converting int types to factor, where appropriate:
```{r}
Allv1 <- FullSetB4
Allv1$DM <- as.factor(Allv1$DM)
Allv1$HTN <- as.factor(Allv1$HTN)
Allv1$Smoker <- as.factor(Allv1$Smoker)
Allv1$ExSmoker <- as.factor(Allv1$ExSmoker)
Allv1$FH <- as.factor(Allv1$FH)
Allv1$Edema <- as.factor(Allv1$Edema)
Allv1$TCP <- as.factor(Allv1$TCP)
Allv1$QWave <- as.factor(Allv1$QWave)
Allv1$STelev <- as.factor(Allv1$STelev)
Allv1$STdep <- as.factor(Allv1$STdep)
Allv1$Tinv <- as.factor(Allv1$Tinv)
#str(Allv1)
```
# DISCRETIZING:

# Change age attribute to factor: If male and age is <= to 45, OR if female and age is <= 55; categorize as "med" (for 'medium').  If male and age is > 45 or female and age is > 55, categorize as "high".
```{r}
Allv1$Age <- factor(ifelse(Allv1$Age <= 45 & Allv1$Sex == 'Male' | Allv1$Age <= 55 & Allv1$Sex == 'Fmale', "Med", "High"))
#str(Allv1)
```
# Checking for any errors in Age factor values, and the proportion of 'medium' to 'high' ages (after being discretized):
```{r}
MedAge <- sum(Allv1$Age == 'Med')
HighAge <- sum(Allv1$Age == 'High')
sum(MedAge+HighAge)
(MedAge/(MedAge+HighAge))*100
```
# Discretize Weight for males vs females. These discretized ranges were not specified in the original study, so I am using the mean and the factoring into 'below'("Ave") or 'above' ("High") the group average:
```{r}
tapply(Allv1$Weight, Allv1$Sex, mean)

Allv1$Weight <- factor(ifelse(Allv1$Weight <= 76.2 & Allv1$Sex == 'Male' | Allv1$Weight <= 70.5 & Allv1$Sex == 'Fmale', "Ave", "High"))
```
# Discretize Blood Pressure (BP):
```{r}
max(Allv1$BP)
Allv1$BP <- cut(Allv1$BP, breaks = c(0,89,140,200), labels = c("Low", "Med", "High"))

BPlow <- sum(Allv1$BP == "Low")
BPlow
BPmed <- sum(Allv1$BP == "Med")
BPmed
BPhigh <- sum(Allv1$BP == "High")
BPhigh
BPcount <- sum(BPlow+BPmed+BPhigh)
BPcount
```
# Note that no patients had a 'low' blood pressure. Check the proportion of medium to high blood pressure results:
```{r}
(BPmed/(BPmed+BPhigh))*100
```
# Discretize PUlse Rate (PR). Note: Very low variation in pulse rate (98.3% was categorized with a 'medium' heart rate). THis would make the 'Pulse rate' attribute a good candidate for removal.
```{r}
max(Allv1$PR)
Allv1$PR <- cut(Allv1$PR, breaks = c(0,59,100,200), labels = c("Low", "Med", "High"))

PRlow <- sum(Allv1$PR == "Low")
PRlow
PRmed <- sum(Allv1$PR == "Med")
PRmed
PRhigh <- sum(Allv1$PR == "High")
PRhigh
PRcount <- sum(PRlow+PRmed+PRhigh)
PRcount
```
# Discretize Heart Rate Functional Class (Fclass): Class 0 equals 'Med' and classes 1, 2, and 3 equal 'High'.  Note that 69.6% of the patients were evaluated to have a "Medium" Heart Failure Functional Class.
```{r}
sum(Allv1$Fclass == 0)
Allv1$Fclass <- ifelse(Allv1$Fclass == 0, "Med", "High")

FCmed <- sum(Allv1$Fclass == "Med")
FCmed
FChigh <- sum(Allv1$Fclass == "High")
FChigh
sum(FCmed+FChigh)
(FCmed/(FCmed+FChigh))*100
```
# Count the levels and results of the 'Exertional Chest Pain' (ECP) attribute (yes/no entries).  
```{r}
nlevels(Allv1$ECP)
sum(Allv1$ECP == 'N')
```
# Deleting this attribute given that there is no variation in the results, and as such, does not improve predictive effectiveness:
```{r}
Allv1$ECP <- NULL
Allv2 <- Allv1
```


# Checking 'Bundle Branch Block' (BBB) for the numbers of LBBB and RBBB results there are in the total (93% of patients had no BBB).
```{r}
sum(Allv2$BBB == 'N')
sum(Allv2$BBB == 'LBBB')
sum(Allv2$BBB == 'RBBB')

```

# Discretizing the 'Fasting Blood Sugar' attribute into 'low', 'med' and 'high':
```{r}
max(Allv2$FBS)
Allv2$FBS <- cut(Allv2$FBS, breaks = c(0,69,105,500), labels = c("Low", "Med", "High"))

FBSlow <- sum(Allv2$FBS == "Low")
FBSlow
FBSmed <- sum(Allv2$FBS == "Med")
FBSmed
FBShigh <- sum(Allv2$FBS == "High")
FBShigh
FBScount <- sum(FBSlow+FBSmed+FBShigh)
FBScount
```
# Discretizing the 'Creatine' attribute into 'low', 'med' and 'high':
```{r}
max(Allv2$Cr)
Allv2$Cr <- cut(Allv2$Cr, breaks = c(0,0.69,1.5,2.5), labels = c("Low", "Med", "High"))

Crlow <- sum(Allv2$Cr == "Low")
Crlow
Crmed <- sum(Allv2$Cr == "Med")
Crmed
Crhigh <- sum(Allv2$Cr == "High")
Crhigh
CrCount <- sum(Crlow+Crmed+Crhigh)
CrCount
```
# Discretizing the Triglyceride (TG) variable into 'med' and 'high':
```{r}
max(Allv2$TG)
Allv2$TG <- cut(Allv2$TG, breaks = c(0,200,1100), labels = c("Med", "High"))

TGmed <- sum(Allv2$TG == "Med")
TGmed
TGhigh <- sum(Allv2$TG == "High")
TGhigh
TGcount <- sum(TGmed+TGhigh)
TGcount
```
# Discretizing Low Density Lipoprotein (LDL). It appears that the results of the cateorization of Triglycerides exactly matches that of LDL, hence, one of these two attributes can be removed.
```{r}
max(Allv2$LDL)
Allv2$LDL <- cut(Allv2$LDL, breaks = c(0,130,250), labels = c("Med", "High"))

LDLmed <- sum(Allv2$LDL == "Med")
LDLmed
LDLhigh <- sum(Allv2$LDL == "High")
LDLhigh
LDLcount <- sum(LDLmed+LDLhigh)
LDLcount
```
# Discretizing High Density Lipoprotein (HDL).
```{r}
max(Allv2$HDL)
Allv2$HDL <- cut(Allv2$HDL, breaks = c(0,34,120), labels = c("Low", "Med"))

HDLlow <- sum(Allv2$HDL == "Low")
HDLlow
HDLmed <- sum(Allv2$HDL == "Med")
HDLmed
HDLcount <- sum(HDLlow+HDLmed)
HDLcount
```
# Discretizing Blood Urea Nitrogen (BUN):
```{r}
max(Allv2$BUN)
Allv2$BUN <- cut(Allv2$BUN, breaks = c(0,6,20,55), labels = c("Low", "Med", "High"))

BUNlow <- sum(Allv2$BUN == "Low")
BUNlow
BUNmed <- sum(Allv2$BUN == "Med")
BUNmed
BUNhigh <- sum(Allv2$BUN == "High")
BUNhigh
BUNcount <- sum(BUNlow+BUNmed+BUNhigh)
BUNcount
```

# Taking steps to discretize 'Erythrocyte Sedimentation Rate' (ESR). If male and ESR <= age/2 OR if female and ESR <= (age/2)+5 then ESR is considered "Mediium".  If male and ESR is > age/2 OR if female and ESR > (age/2)+5, then ESR is considered "High":
```{r}
tempAge <- data.frame(FullSetB4[,1:3, 44])

tempAge$Age <- ifelse(tempAge$Sex == 'Male', tempAge$Age/2, tempAge$Age)
tempAge2 <- tempAge

tempAge2$Age <- ifelse(tempAge2$Sex == 'Fmale', (tempAge2$Age/2)+5, tempAge2$Age)

tempAge2$ESR <- FullSetB4$ESR
tempAge2$Weight <- NULL
tempAge3 <- tempAge2

tempAge3$ESR <- factor(ifelse(tempAge3$ESR <= tempAge3$Age & tempAge3$Sex == 'Male' | tempAge3$ESR <= tempAge3$Age & tempAge3$Sex == 'Fmale', "Med", "High"))
str(tempAge3)

Allv2$ESR <- tempAge3$ESR
```

# Discretizing Hemoglobin (Hb). If male & Hb < 14 or if female and Hb is < 12.5, then Hb is "Low". If male & Hb is >= 14 but <= 17 OR if female and Hb is >= 12 but <= 15, then Hb is "Medium". If male & Hb > 17 or if female and Hb is > 15, then Hb is "High".
```{r}
tempHb <- data.frame(Allv2[,3:44])
tempHb[,2:41] <- NULL

tempHb$Hb <- with(tempHb, ifelse(tempHb$Sex == 'Male' & tempHb$Hb < 14 | tempHb$Sex == 'Fmale' & tempHb$Hb < 12.5, "Low", ifelse(tempHb$Sex == 'Male' & tempHb$Hb > 17 | tempHb$Sex == 'Fmale' & tempHb$Hb > 15, "High", "Med")))

Allv2$Hb <- as.factor(tempHb$Hb)
str(Allv2)
```

# Discretizing Potassium (K):
```{r}
max(Allv2$K)
Allv2$K <- cut(Allv2$K, breaks = c(0,3.7,5.6,6.8), labels = c("Low", "Med", "High"))

Klow <- sum(Allv2$K == "Low")
Klow
Kmed <- sum(Allv2$K == "Med")
Kmed
Khigh <- sum(Allv2$K == "High")
Khigh
Kcount <- sum(Klow+Kmed+Khigh)
Kcount
```

# Discretizing Sodium (Na):
```{r}
max(Allv2$Na)
Allv2$Na <- cut(Allv2$Na, breaks = c(0,135,146,157), labels = c("Low", "Med", "High"))

NaLow <- sum(Allv2$Na == "Low")
NaLow
NaMed <- sum(Allv2$Na == "Med")
NaMed
NaHigh <- sum(Allv2$Na == "High")
NaHigh
NaCount <- sum(NaLow+NaMed+NaHigh)
NaCount
```
# Discretizing White Blood Cell (WBC) count:
```{r}
max(Allv2$WBC)
Allv2$WBC <- cut(Allv2$WBC, breaks = c(0,3999,11000,19000), labels = c("Low", "Med", "High"))

WBClow <- sum(Allv2$WBC == "Low")
WBClow
WBCmed <- sum(Allv2$WBC == "Med")
WBCmed
WBChigh <- sum(Allv2$WBC == "High")
WBChigh
WBCcount <- sum(WBClow+WBCmed+WBChigh)
WBCcount
```
# Discretizing Lymphocyte (Lymph) percentage results (See reference 21 on Literature Review).  Spreading the results into 3 sections: 'Low', 'Med' and 'High'.
```{r}
max(Allv2$Lymph)
Allv2$Lymph <- cut(Allv2$Lymph, breaks = c(0,17,45,65), labels = c("Low", "Med", "High"))

LymphLow <- sum(Allv2$Lymph == "Low")
LymphLow
LymphMed <- sum(Allv2$Lymph == "Med")
LymphMed
LymphHigh <- sum(Allv2$Lymph == "High")
LymphHigh
LymphCount <- sum(LymphLow+LymphMed+LymphHigh)
LymphCount
```
# Discretizing Neutrophil (Neut) percentage results (See reference 22 on Literature Review).  Spreading the results into 3 sections: 'Low', 'Med' and 'High'.
```{r}
max(Allv2$Neut)
Allv2$Neut <- cut(Allv2$Neut, breaks = c(0,44,75,90), labels = c("Low", "Med", "High"))

NeutLow <- sum(Allv2$Neut == "Low")
NeutLow
NeutMed <- sum(Allv2$Neut == "Med")
NeutMed
NeutHigh <- sum(Allv2$Neut == "High")
NeutHigh
NeutCount <- sum(NeutLow+NeutMed+NeutHigh)
NeutCount
```
# Discretizing Ejection Fraction (EF):
```{r}
max(Allv2$EF)
Allv2$EF <- cut(Allv2$EF, breaks = c(0,50,120), labels = c("Low", "Med"))

EFlow <- sum(Allv2$EF == "Low")
EFlow
EFmed <- sum(Allv2$EF == "Med")
EFmed
EFcount <- sum(EFlow+EFmed)
EFcount
```
# Discretizing Platelet (PLT) count (Note: 96% of patients had 'Medium' counts)
```{r}
max(Allv2$PLT)
Allv2$PLT <- cut(Allv2$PLT, breaks = c(0,149,450,750), labels = c("Low", "Med", "High"))

PLTlow <- sum(Allv2$PLT == "Low")
PLTlow
PLTmed <- sum(Allv2$PLT == "Med")
PLTmed
PLThigh <- sum(Allv2$PLT == "High")
PLThigh
PLTcount <- sum(PLTlow+PLTmed+PLThigh)
PLTcount
```
# Discretizing Regional Wall Motion Abnormality (RWMA). RWMA = 0 is 'Med', RWMA != to 0 are 'High').  Found that 71.6% were found to be 'Medium'.
```{r}
sum(Allv2$RWMA == 0)
Allv2$RWMA <- ifelse(Allv2$RWMA == 0, "Med", "High")

RWMAmed <- sum(Allv2$RWMA == "Med")
RWMAmed
RWMAhigh <- sum(Allv2$RWMA == "High")
RWMAhigh
sum(RWMAmed+RWMAhigh)
(RWMAmed/(RWMAmed+RWMAhigh))*100
```
# Removing the 'BMI' attribute since this is captured in the 'Obesity' attribute (given that a BMI result that is over 25 is considered 'Obese'). Matching the removed attributes for the 'before' and 'after' (discretization) data sets:
```{r}
Allv2$BMI <- NULL

FullSetB4$BMI <- NULL
FullSetB4$ECP <- NULL
```

# Taking the completed (and discretized) 'Allv2' data frame and saving it as 4 separate attribute groups (data frames); to be used for the rest of the predictions/modeling.  Keeping the dependent result (CAD) at this stage:
```{r}
Demo <- data.frame(Allv2[,1:15])
Demo$CAD <- Allv2$CAD

Symptom <- data.frame(Allv2[,16:28])
Symptom$CAD <- Allv2$CAD

ECG <- data.frame(Allv2[,29:35])
ECG$CAD <- Allv2$CAD

LAB <- data.frame(Allv2[,36:52])
LAB$CAD <- Allv2$CAD
```

# Taking the data set before discretizing and factoring (FullSetB4) and saving it as 4 separate attribute groups (data frames); to be used for analysis. First removing the 'ECP' and 'BMI' attributes (to match the discretized data set):
```{r}
FullSetB42 <- FullSetB4

DemoB4 <- data.frame(FullSetB42[,1:15])
DemoB4$CAD <- FullSetB42$CAD

SymptomB4 <- data.frame(FullSetB42[,16:28])
SymptomB4$CAD <- FullSetB42$CAD

ECGB4 <- data.frame(FullSetB42[,29:35])
ECGB4$CAD <- FullSetB42$CAD

LABB4 <- data.frame(FullSetB42[,36:52])
LABB4$CAD <- FullSetB42$CAD
```

UNIVARIATE ANALYSIS:

#  Correlation coefficients for the 'LAB' results show no correlation (other than the negative correlation between Lymphocyte and Neutrophil results). With regards to this inverse relationship, an abnormal increase in one kind of white blood cell can cause a decrease in another kind. Both abnormal results can be due to the same underlying condition.
```{r}
library(GGally)

LABcor3 <- ggcorr(LABB4[1:15], palette = "RdY1Gn", label = TRUE)
LABcor3
```
# Looking into the Neutrophil to Lymphocyte relationship further.  Calculating the NLR (neutrophil-lymphocyte ratio) below. Reference # 24 shows that they have identified that the normal NLR values in an adult, non-geriatric, population in good health are between 0.78 and 3.53. 
```{r}
NLR <- data.frame(FullSetB42[48:47])
NLR$Ratio <- NLR$Neut/NLR$Lymph

min(NLR$Ratio)
max(NLR$Ratio)
mean(NLR$Ratio)

NLR$Result <- cut(NLR$Ratio, breaks = c(0,0.77,3.53,12.5), labels = c("Low", "Med", "High"))

NLR$CAD <- FullSetB42$CAD

NLRlow <- sum(NLR$Result == "Low")
NLRlow
NLRmed <- sum(NLR$Result == "Med")
NLRmed
NLRhigh <- sum(NLR$Result == "High")
NLRhigh
NLRcount <- sum(NLRlow+NLRmed+NLRhigh)
NLRcount
```

# Include remaining numeric attributes for further correlation overview: Shows no significant correlation results, other than what was already found:
```{r}
library(GGally)

Allints1 <- ggcorr(FullSetB42[,c(1:2, 16:17, 36:50)], palette = "RdY1Gn", label = TRUE)
Allints1
```
# Boxplot of Low Density Lipoprotein (LDL), High Density Lipoprotein (HDL), Blood Urea Nitrogen (BUN), and Erythrocyte Sedimentation Rate (ESR):
```{r}
attach(LABB4)
boxplot(LDL, HDL, BUN, ESR, col = c("cadetblue1", "antiquewhite", "chartreuse", "darkseagreen1"), notch = TRUE, horizontal = FALSE, outline = TRUE, plot = TRUE)
legend("topright", c("LDL", "HDL", "BUN", "ESR"),fill = c("cadetblue1", "antiquewhite", "chartreuse", "darkseagreen1"))
```
# Boxplot of Sodium (Na), Lymphocyte % (Lymph), Neutrophil % (Neut), and Ejection Fraction % (EF):
```{r}
boxplot(Na, Lymph, Neut, EF, col = c("indianred1", "antiquewhite", "chartreuse", "gold"), notch = TRUE, horizontal = FALSE, outline = TRUE, plot = TRUE)
legend("topright", c("Na", "Lymph", "Neut", "EF"),fill = c("indianred1", "antiquewhite", "chartreuse", "gold"))
```
# Boxplot of Triglyceride (TG), Platelet (PLT), and Fasting Blood Sugar (FBS):
```{r}
boxplot(TG, PLT, FBS, col = c("aquamarine","indianred1", "gray90"), notch = TRUE, horizontal = FALSE, outline = TRUE, plot = TRUE)
legend("topright", c("TG", "PLT", "FBS"),fill = c("aquamarine","indianred1", "gray90"))
```
# Boxplot of Hemoglobin (Hb), Potassium (K), and Creatine (Cr):
```{r}
boxplot(Hb, K, Cr, col = c("aquamarine","indianred1", "gray90"), notch = TRUE, horizontal = FALSE, outline = TRUE, plot = TRUE)
legend("topright", c("Hb", "K", "Cr"),fill = c("aquamarine","indianred1", "gray90"))
```
# Boxplot of White Blood Cell (WBC) count:
```{r}
boxplot(WBC, col = c("antiquewhite"), notch = TRUE, horizontal = FALSE, outline = TRUE, plot = TRUE)
legend("topright", c("WBC"),fill = c("antiquewhite"))
```
# Boxplot to show distribution of Age and Weight:
```{r}
attach(DemoB4)
boxplot(Age, Weight, col = c("aquamarine","gray90"), notch = TRUE, horizontal = FALSE, outline = TRUE, plot = TRUE)
legend("topright", c("Age", "Weight"),fill = c("aquamarine","gray90"))
```
# Shapiro tests of normality (H0: Distribution is normal; Reject null if p-value is less than .05). Note: sensitive when n>80. This test shows that 'Age' is NOT normally distributed.
```{r}
library(psych)

describe(DemoB4$Age)
AgeNorm <- shapiro.test((DemoB4$Age))
AgeNorm
```
# Hence, weight is also not normally distributed in this study.
```{r}
library(psych)

describe(DemoB4$Weight)
WeightNorm <- shapiro.test((DemoB4$Weight))
WeightNorm
```
# And BMI is also not normally distributed.
```{r}
describe(Allv1$BMI)
BMINorm <- shapiro.test((Allv1$BMI))
BMINorm
```
# Can (carefully) use 'describe' for categorical variables too (since the psych package recodes categories as numbers):
```{r}
library(psych)
describe(FullSetB42[1:17])
```
# Boxplot to show distribution of Blood Pressure and Pulse Rate
```{r}
attach(SymptomB4)
boxplot(BP, PR, col = c("blue","indianred1"), notch = TRUE, horizontal = FALSE, outline = TRUE, plot = TRUE)
legend("topright", c("Blood Pressure", "Pulse Rate"),fill = c("blue","indianred1"))
```
# Blood Pressure (BP) is not normally distributed:
```{r}
describe(SymptomB4$BP)
BPNorm <- shapiro.test((SymptomB4$BP))
BPNorm
```
# Pulse Rate is not normally distributed:
```{r}
describe(SymptomB4$PR)
PRNorm <- shapiro.test((SymptomB4$PR))
PRNorm
```
# Density plot of weight and sex (before discretization):
```{r}
library(ggplot2)
library(dplyr)
library(plyr)

WeightMu <- ddply(DemoB4, "Sex", summarise, grp.mean=mean(Weight))
head(WeightMu)

DDen1 <- ggplot(DemoB4, aes(x = Weight, fill = Sex)) +
  geom_density(alpha = 0.2)+
  geom_vline(data = WeightMu, aes(xintercept=grp.mean, colour = Sex),
             linetype = "dashed")
DDen1
```
# Density plot of BMI and sex (before discretization):
```{r}
BMIMu <- ddply(Allv1, "Sex", summarise, grp.mean=mean(BMI))
head(BMIMu)

BMIden <- ggplot(Allv1, aes(x = BMI, fill = Sex)) +
  geom_density(alpha = 0.2)+
  geom_vline(data = BMIMu, aes(xintercept=grp.mean, colour = Sex),
             linetype = "dashed")
BMIden
```
# Will be completing a number of ASSOCIATIONS between many categorical variables, starting with a few straight forward investigation into commonly accepted 'associations'. Interpretation of results when K = 2:  SMALL Association: 0.10 - < 0.30; MEDIUM Association: 0.30 - < 0.50; LARGE Association: ??? 0.50

# Starting with Diabetes (DM) and Hypertension (HTN): Shows small association
```{r}
library(lsr)

AllCatB4 <- Allv2[,c(3:15, 18:24,26:35)]
cramersV(AllCatB4$DM, AllCatB4$HTN)
```
# Ex-Smoker and Obesity shows no association:
```{r}
cramersV(AllCatB4$ExSmoker,AllCatB4$Obesity)
```
# Ex-Smoker and Obesity shows no association:
```{r}
cramersV(AllCatB4$DM,AllCatB4$Obesity)
```

# Hypertensin and current smoker has a small association:
```{r}
cramersV(AllCatB4$HTN,AllCatB4$Smoker)
```

# Hypertension and Stroke (CVA) has no association:
```{r}
cramersV(AllCatB4$HTN,AllCatB4$CVA)
```
# Congestive Heart Failure (CHF) and Obsesity has no association (see below):
```{r}
cramersV(AllCatB4$CHF,AllCatB4$Obesity)
```

# Determining whether there is any significant variation in the 'NCP' column: Very low variation (94.7% of NCP entries are 'No')
```{r}
NCPn <- sum(FullSetB42$NCP == 'N')
NCPy <- sum(FullSetB42$NCP == 'Y')
(NCPn/(NCPn+NCPy))*100
```

# Determining whether there is any significant variation in the 'ACP' column: 69.3% of these 'ACP' entries are 'No'
```{r}
ACPn <- sum(FullSetB42$ACP == 'N')
ACPy <- sum(FullSetB42$ACP == 'Y')
(ACPn/(ACPn+ACPy))*100
```

# Barplot of BBB attribute
```{r}
barplot(prop.table(table(FullSetB42$BBB)))
```
# Barplot of VHD attribute
```{r}
barplot(prop.table(table(FullSetB42$VHD)))
```

# Feature selection on all the attributes combined, as well as on the 4 individual attribute groups - using the FSelector package:

# Starting with the full data set - selected: CAD ~ TCP + ACP + RWMA + EF + Age + HTN + NCP + DM + Tinv + VHD
```{r}
library(FSelector)

AllWeights <- chi.squared(CAD~.,Allv2)
print(AllWeights)
subset <- cutoff.k(AllWeights, 10)
FSall <- as.simple.formula(subset, "CAD")
print(FSall)
```

# Next, using the 'Demo' (Demographic) attribute data set - results:  CAD ~ Age + HTN + DM + CRF + AD
```{r}
DemoWeights <- chi.squared(CAD~.,Demo)
print(DemoWeights)
subset <- cutoff.k(DemoWeights, 5)
FSdemo <- as.simple.formula(subset, "CAD")
print(FSdemo)
```

# Next, using the 'Symptom' (Symptom & Examination) attribute data set - results:  CAD ~ TCP + ACP + NCP + DiaM + BP
```{r}
SymWeights <- chi.squared(CAD~.,Symptom)
print(SymWeights)
subset <- cutoff.k(SymWeights, 5)
FSsym <- as.simple.formula(subset, "CAD")
print(FSsym)
```

# Next, using the 'ECG'attribute data set - results:  CAD ~ Tinv + QWave + STdep + STelev + PoorR
```{r}
ECGWeights <- chi.squared(CAD~.,ECG)
print(ECGWeights)
subset <- cutoff.k(ECGWeights, 5)
FSecg <- as.simple.formula(subset, "CAD")
print(FSecg)
```

# Lastly, using the 'LAB' (Laboratory and Echocardiographic) attribute data set - results:  CAD ~ RWMA + EF + VHD + FBS + Lymph 
```{r}
LABWeights <- chi.squared(CAD~.,LAB)
print(LABWeights)
subset <- cutoff.k(LABWeights, 5)
FSlab <- as.simple.formula(subset, "CAD")
print(FSlab)
```

# Now, comparing these discretized results with those from before discretization:

# Starting with the full data set - results: CAD ~ TCP + ACP + RWMA + EF + Age + HTN + *BP* + NCP + DM + *FBS* (while the discretized results were: CAD ~ TCP + ACP + RWMA + EF + Age + HTN + NCP + DM + Tinv + VHD).  NOTE: The top 6 were the same for both data sets (and 8 of the top 10 were shared across both data sets)
```{r}
AllB4Weights <- chi.squared(CAD~.,FullSetB42)
print(AllB4Weights)
subset <- cutoff.k(AllB4Weights, 10)
FSB4all <- as.simple.formula(subset, "CAD")
print(FSB4all)
```

# Comparing the discretized feature selection results found using the 'FSelector' package with the results using the 'FSelectorRcpp' package - results: CAD ~ TCP + ACP + RWMA + EF + HTN + Age + DM + NCP + Tinv + FBS
```{r}
library(FSelectorRcpp)
library(tidyverse)

FS2all <- information_gain(
  formula = CAD~.,
  data = Allv2,
  type = "infogain",
  threads = 2
) %>%
cut_attrs(
  k = 10
) %>%
to_formula(
  attrs = .,
  class = "CAD"
)
print(FS2all)

```

# Lastly, comparing the output of the FSelectorRcpp package using the data set that was not discretized yet (FullSetB42 data set) - result:  CAD ~ TCP + ACP + RWMA + Age + EF + HTN + DM + BP + NCP + Tinv
```{r}
FS2B4all <- information_gain(
  formula = CAD~.,
  data = FullSetB42,
  type = "infogain",
  threads = 2
) %>%
cut_attrs(
  k = 10
) %>%
to_formula(
  attrs = .,
  class = "CAD"
)
print(FS2B4all)
```



